---
title: "Study in figure properties"
output:
  pdf_document:
    fig_crop: no
    highlight: kate
    number_sections: yes
    toc: yes
    toc_depth: 3
  md_document:
    toc: yes
    variant: markdown
  html_document:
    css: ~/GitHub/Longitudinal_Models_of_Religiosity_NLSY97/www/css/thesis.css
    fig_caption: yes
    fig_height: 4.8
    fig_width: 6.5
    highlight: textmate
    keep_md: yes
    number_sections: yes
    theme: united
    toc: yes
---

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
require(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.

```

```{r set_options, echo=F, message=F}
require(knitr)
# getwd()
opts_chunk$set(
  results='show', 
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
  fig.height = 4.8, 
  fig.width = 6.5, 
  out.width = "550px",
  fig.path = 'figure_rmd/',     
  dev = "png",
#   fig.path = 'figure_pdf/',     
#   dev = "pdf",
  dpi = 400
)
echoChunks <- FALSE
warningChunks<- FALSE
options(width=120) #So the output is 50% wider than the default.
```
Exploring the figure definitions

+ fig.width
+ fig.height
+ out.width

```{r DeclareGlobals, echo=echoChunks, message=FALSE}
attcol8<-c("Never"="#4575b4",
           "Once or Twice"="#74add1",
           "Less than once/month"="#abd9e9",
           "About once/month"="#e0f3f8",
           "About twice/month"="#fee090",
           "About once/week"="#fdae61",
           "Several times/week"="#f46d43",
           "Everyday"="#d73027")
```

```{r LoadPackages, echo=echoChunks, message=F}
require(dplyr)
require(ggplot2)
```

```{r LoadData, echo=echoChunks, message=T}
dsL<-readRDS("./Data/Derived/dsL.rds")
```


```{r DefineCommonGraph, echo=echoChunks, message=T}
p<-ggplot(subset(dsL,year==2000), aes(x=attendF, fill=attendF))
p <- p + geom_bar(color="black")
p <- p + scale_fill_manual(values=attcol8, na.value="white")
p <- p + coord_flip()
p <- p + labs(title="How often did you attend a place of worship in the last year? (2000)", x="Church attendance", y="Count")
```


# Fig.hight

## fig.height = 4, fig.width=7, out.width = "600px"
```{r 4_7_600, echo=echoChunks, message=T, fig.height = 4, fig.width=7, out.width = "600px"}
p
```


## fig.height = 3.5, fig.width=7, out.width = "600px"
```{r 3.5_7_600, echo=echoChunks, message=T, fig.height = 3.5, fig.width=7, out.width = "600px"}
p
```

## fig.height = 3, fig.width=7, out.width = "600px"
```{r 3_7_600, echo=echoChunks, message=T, fig.height = 3, fig.width=7, out.width = "600px"}
p
```


## fig.height = 2.5, fig.width=7, out.width = "600px"
```{r 2.5_7_600, echo=echoChunks, message=T, fig.height = 2.5, fig.width=7, out.width = "600px"}
p
```


## fig.height = 2, fig.width=7, out.width = "600px"
```{r 2_7_600, echo=echoChunks, message=T, fig.height = 2, fig.width=7, out.width = "600px"}
p
```


# Fig.width

## fig.height = 4, fig.width=7, out.width = "600px"
```{r 4_7_600_2, echo=echoChunks, message=T, fig.height = 4, fig.width=7, out.width = "600px"}
p
```


## fig.height = 4, fig.width=6.5, out.width = "600px"
```{r 4_6.5_600, echo=echoChunks, message=T, fig.height = 4, fig.width=6.5, out.width = "600px"}
p
```


## fig.height = 4, fig.width=6, out.width = "600px"
```{r 4_6_600, echo=echoChunks, message=T, fig.height = 4, fig.width=6, out.width = "600px"}
p
```


## fig.height = 4, fig.width=5.5, out.width = "600px"
```{r 4_5.5_600, echo=echoChunks, message=T, fig.height = 4, fig.width=5.5, out.width = "600px"}
p`


## fig.height = 4, fig.width=5, out.width = "600px"
```{r 4_5_600, echo=echoChunks, message=T, fig.height = 4, fig.width=5, out.width = "600px"}
p
```


## fig.height = 4, fig.width=4.5, out.width = "600px"
```{r 4_3.5_600, echo=echoChunks, message=T, fig.height = 4, fig.width=4.5, out.width = "600px"}
p
```


# out.width

## fig.height = 4, fig.width=7, out.width = "600px"
```{r 4_7_600_3, echo=echoChunks, message=T, fig.height = 4, fig.width=7, out.width = "600px"}
p
```


## fig.height = 4, fig.width=7, out.width = "550px"
```{r 4_7_550, echo=echoChunks, message=T, fig.height = 4, fig.width=7, out.width = "550px"}
p
```

## fig.height = 4, fig.width=7, out.width = "500px"
```{r 4_7_500, echo=echoChunks, message=T, fig.height = 4, fig.width=7, out.width = "500px"}
p
```


## fig.height = 4, fig.width=7, out.width = "450px"
```{r 4_7_450, echo=echoChunks, message=T, fig.height = 4, fig.width=7, out.width = "450px"}
p
```


## fig.height = 4, fig.width=7, out.width = "400px"
```{r 4_7_400, echo=echoChunks, message=T, fig.height = 4, fig.width=7, out.width = "400px"}
p
```



## fig.height = 4, fig.width=7, out.width = "350px"
```{r 4_7_350, echo=echoChunks, message=T, fig.height = 4, fig.width=7, out.width = "350px"}
p
```
