---
title: "Data Manipulation"
output:
  html_document:
    css: ~/GitHub/Longitudinal_Models_of_Religiosity_NLSY97/www/css/thesis.css
    fig.retina: 2
    fig_width: 8
    toc: yes
  pdf_document:
    fig_width: 8
    latex_engine: xelatex
    toc: yes
  word_document:
    fig_width: 6.5
mainfont: Calibri
---

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
require(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.

```

```{r set_options, echo=F, message=F}
require(knitr)
# getwd()
opts_chunk$set(
  results='show', 
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
#   fig.height = 4.8, 
#   fig.width = 6.5, 
#   out.width = "550px",
  fig.path = 'figure_rmd/automatic_report/',     
  dev = "png",
#   fig.path = 'figure_pdf/',     
#   dev = "pdf",
  dpi = 400
)
echoChunks <- FALSE
warningChunks<- FALSE
options(width=120) #So the output is 50% wider than the default.
read_chunk("./Vignettes/dplyr/using_dplyr.R") # the file to which knitr calls for the chunks
```

Data Manipulation
==============
Report examplifying the use of <code>dplyr</code> in data handling on the example of **dsL**>.

<!-- Run this three chunks to get to the starting point -->
```{r DeclareGlobals, echo=echoChunks, message=FALSE}
```

```{r LoadPackages, echo=echoChunks, message=F}
```

```{r LoadData, echo=echoChunks, message=FALSE}
```

## Five basic functions in data handling

For a more detailed discussion of basic verbs and operations consult the [R-Studio guide][1] or internal [vignette][2]
```{r,eval=FALSE}
vignette("introduction",package="dplyr")
```

The following is a brief demonstration of <code>dplyr</code> syntax using **dsL** dataset as an example. I attach prefix <code>dplyr::</code> to avoid possible conflicts with <code>plyr</code> package on which <code>ggplot2</code> package relies. I recommend such practice in all <code>dplyr</code>  expressions in sharable publications.  

###<code>select()</code> 
selects variables into a smaller data set
```{r}
ds<-dsL
dim(ds)
ds<- dplyr::select(ds,id,year, byear, attend, attendF)
head(ds,13)
dim(ds)
```

###<code>filter()</code> 
Removes observations that do not meet criteria. The following code selects observation based on the type of sample    

```{r, echo=F}
ds<- dplyr::filter(dsL, id %in% c(1, 6803) & year==2000 )
ds<- dplyr::select(ds, sample, sampleF)
ds
```
and only between years 2000 and 2011, as only during those years the outcome of interest <code>attend</code> was recorded. 
```{r, echo=T, message=F}
ds<- dplyr::filter(dsL,sample==1, year %in% c(2000:2011))
ds<- dplyr::select(ds,id, year, attend, attendF)
head(ds,13)
```


###<code>arrange()</code> 
Sorts observations
```{r, echo=T, message=F}
ds<- dplyr::filter(dsL,sample==1, year %in% c(2000:2011))
ds<- dplyr::select(ds,id, year, attend)
ds<- dplyr::arrange(ds, year, desc(id))
head(ds,13)
ds<- arrange(ds, id, year)
head(ds, 13)
```


###<code>mutate()</code> 

Creates additional variables from the values of existing.
```{r, echo=T, message=F}
ds<- dplyr::filter(dsL,sample==1, year %in% c(2000:2011))
ds<- dplyr::select(ds,id, byear, year, attend)
ds<- dplyr::mutate(ds, 
            age = year-byear, 
            timec = year-2000,
            linear= timec,
            quadratic= linear^2,
            cubic= linear^3)
head(ds,13)
```

###<code>summarize()</code> 
collapses data into a single value computed according to the aggregate functions.


```{r}
require(dplyr)
ds<- dplyr::filter(dsL,sample==1)
ds<- dplyr::summarize(ds, N= n_distinct(id))
ds
```
Other functions one could use with <code>summarize()</code> include:

From <code>base</code>
+ <code> min() </code>   
+ <code> max() </code>   
+ <code> mean() </code>   
+ <code> sum() </code>   
+ <code> sd() </code>   
+ <code> median() </code>   
+ <code> IQR()  </code>   

Native to <code>dplyr</code>
+ <code> n() </code>  -  number of observations in the current group
+ <code> n_distinct(x) </code>  -    count the number of unique values in x.
+ <code> first(x) </code>  -  similar to <code>x[1]</code> + control over <code>NA</code>
+ <code> last(x) </code>  -  similar to <code>x[length(x)] </code> + control over <code>NA</code>
+ <code> nth(x, n)  </code>  -  similar to<code> x[n] </code>  + control over <code>NA</code>

## Grouping and Combining 
The function <code>group_by()</code> is used to identify groups in split-apply-combine (SAC) procedure: it splits the initial data into smaller datasets (according to all possible interactions between the levels of supplied variables). It is these smaller datasets that <code>summarize()</code> will individually collapse into a single computed value according to its formula.  
```{r, echo=T, message=F}
ds<- dplyr::filter(dsL,sample==1, year %in% c(2000:2011))
ds<- dplyr::select(ds,id, year, attendF)

s <- dplyr::group_by(ds, year,attendF)
s <- dplyr::summarise(s, count = n())
s <- dplyr::mutate(s, total = sum(count),
              percent= count/total)
head(s,10)

```

The same result can be achieved with a more elegant syntax that relies on <code>%>%</code> operator, in which <code>x %>% f(y)</code> turns into <code>f(x, y) </code>. Alternatively, one can use <code>%.%</code> for identical results.

```{r, echo=TRUE}
ds<-dsL %>%
  dplyr::filter(sample==1, year %in% c(2000:2011)) %>%
  dplyr::select(id, year, attendF) %>%
  dplyr::group_by(year,attendF) %>%
    dplyr::summarise(count = n()) %>%
    dplyr::mutate(total = sum(count),
              percent= count/total)  
head(ds,10)    
```

To verify that this is what we wanted to achieve:
```{r, echo=T, message=F}
dplyr::summarize( filter(s, year==2000), should.be.one=sum(percent))
```


## Base subsetting
Generally, we can compose any desired dataset by using matrix calls. The general formula is of the form:
**ds**[  _rowCond_  ,  _colCond_  ], where **ds** is a dataframe, and  _rowCond_ and _colCond_ are conditions for including rows and columns of the new dataset, respectively. One can also call a variable by attaching <code> $ </code> followed variable name to the name of the dataset:    <code>**ds**_$variableName_</code>. 
```{r, echo=T}
ds<-dsL[dsL$year %in% c(2000:2011),c('id',"byear","year","attendF","ageyearF","agemon")]
print(ds[ds$id==1,]) 
```

The following is a list of operatiors that can be used in these calls. 
<ul>
<li>basic math operators: <code>+</code>, <code>-</code>, <code>*</code>, <code>/</code>, <code>%%</code>, <code>^</code></li>
<li>math functions: <code>abs</code>, <code>acos</code>, <code>acosh</code>, <code>asin</code>, <code>asinh</code>, <code>atan</code>, <code>atan2</code>,
<code>atanh</code>, <code>ceiling</code>, <code>cos</code>, <code>cosh</code>, <code>cot</code>, <code>coth</code>, <code>exp</code>, <code>floor</code>,
<code>log</code>, <code>log10</code>, <code>round</code>, <code>sign</code>, <code>sin</code>, <code>sinh</code>, <code>sqrt</code>, <code>tan</code>, <code>tanh</code></li>
<li>logical comparisons: <code>&lt;</code>, <code>&lt;=</code>, <code>!=</code>, <code>&gt;=</code>, <code>&gt;</code>, <code>==</code>, <code>%in%</code></li>
<li>boolean operations: <code>&amp;</code>, <code>&amp;&amp;</code>, <code>|</code>, <code>||</code>, <code>!</code>, <code>xor</code></li>
<li>basic aggregations: <code>mean</code>, <code>sum</code>, <code>min</code>, <code>max</code>, <code>sd</code>, <code>var</code></li>
</ul>

<code>dplyr</code> can translate all of these into SQL. For more of on <code>dplyr</code> and SQL compatibility consult another built-in [vignette][3]
```{r,eval=FALSE}
vignette("database",package="dplyr")
```

## Base Reference

The following unary and binary operators are defined for <code> base</code>. They are listed in precedence groups, from highest to lowest.


+ <code> :: :::   </code> -  access variables in a namespace  
+ <code> $ @	    </code> - component / slot extraction   
+ <code> [ [[	    </code> - indexing  
+ <code> ^	      </code>  - exponentiation (right to left)   
+ <code> - +	 </code>  - unary minus and plus   
+ <code> :	 </code>  - sequence operator   
+ <code> %any%	 </code>  - special operators (including %% and %/%)   
+ <code> * /	 </code>  - multiply, divide   
+ <code> + -	 </code>  - (binary) add, subtract   
+ <code> < > <= >= == !=	 </code>  - ordering and comparison   
+ <code> !	 </code>  - negation   
+ <code> & &&	 </code>  - and   
+ <code> | ||	 </code>  - or   
+ <code> ~	 </code>  - as in formulae   
+ <code> -> ->>	 </code>  - rightwards assignment   
+ <code> <- <<-	 </code>  - assignment (right to left)   
+ <code> =	 </code>  - assignment (right to left)   
+ <code> ?	 </code>  - help (unary and binary)   


[1]:http://blog.rstudio.org/2014/01/17/introducing-dplyr/
[2]:http://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html
[3]:http://cran.rstudio.com/web/packages/dplyr/vignettes/databases.html
