---
title: "Raw vs Summarized"
output:
  html_document:
    css: ~/GitHub/Longitudinal_Models_of_Religiosity_NLSY97/www/css/thesis.css
    fig_caption: yes
    fig_height: 4.5
    fig_width: 6.5
    highlight: textmate
    keep_md: yes
    theme: united
    toc: yes
  md_document:
    toc: yes
    variant: markdown
  pdf_document:
    fig_crop: no
    highlight: kate
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 3
---

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
require(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.

```

```{r , echo=F, message=F}
require(knitr)
# getwd()
opts_chunk$set(
  results='show', 
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
  fig.height = 4.8, 
  fig.width = 6.5, 
  out.width = "550px",
  fig.path = 'figure_rmd/automatic_report/',     
  dev = "png",
#   fig.path = 'figure_pdf/',     
#   dev = "pdf",
  dpi = 400
)
echoChunks <- FALSE
warningChunks<- FALSE
options(width=120) #So the output is 50% wider than the default.
# read_chunk("./Vignettes/dplyr/using_dplyr.R") # the file to which knitr calls for the chunks
```

### Same plot from raw and summarized data

For the following data structure
```{r, message=F}
dsN<-data.frame(
  id=rep(1:100, each=4),
  yearF=factor(rep(2001:2004, 100)),
  attendF=sample(1:8, 400, T, c(.2,.2,.15,.10,.10, .20, .15, .02))
)
dsN[sample(which(dsN$yearF==2001), 5), "attendF"]<-NA
dsN[sample(which(dsN$yearF==2002), 10), "attendF"]<-NA
dsN[sample(which(dsN$yearF==2003), 15), "attendF"]<-NA
dsN[sample(which(dsN$yearF==2004), 20), "attendF"]<-NA

attcol8<-c("Never"="#4575b4",
           "Once or Twice"="#74add1",
           "Less than once/month"="#abd9e9",
           "About once/month"="#e0f3f8",
           "About twice/month"="#fee090",
           "About once/week"="#fdae61",
           "Several times/week"="#f46d43",
           "Everyday"="#d73027")
dsN$attendF<-factor(dsN$attendF, levels=1:8, labels=names(attcol8))
head(dsN,13)
```

we can obtain a series of a stacked bar charts
```{r, message=F}
require(ggplot2)
# p<- ggplot( subset(dsN,!is.na(attendF)), aes(x=yearF, fill=attendF)) # leaving NA out of
p<- ggplot( dsN, aes(x=yearF, fill=attendF))  #  keeping NA in calculations
p<- p+ geom_bar(position="fill")
p<- p+ scale_fill_manual(values = attcol8,
                         name="Response category" )
p<- p+ scale_y_continuous("Prevalence: proportion of total",
                          limits=c(0, 1),
                          breaks=c(.1,.2,.3,.4,.5,.6,.7,.8,.9,1))
p<- p+ scale_x_discrete("Waves of measurement",
                        limits=as.character(c(2000:2005)))
p<- p+ labs(title=paste0("In the past year, how often have you attended a worship service?"))
p
```

The graph above is produced from the raw data. However, it is sometimes convenient to produce graphs from summarized data, especially if one needs control over statistical functions. Below is transformation of dsN into ds that contains only the values that are actually mapped on the graph above:

```{r, message=F}
require(dplyr)
ds<- dsN %.%
  dplyr::filter(!is.na(attendF)) %.%
  dplyr::group_by(yearF,attendF) %.%
  dplyr::summarize(count = sum(attendF)) %.%
  dplyr::mutate(total = sum(count),
              percent= count/total)
head(ds,10)
# verify
summarize(filter(ds, yearF==2001), should.be.one=sum(percent))
```

### Question:
How would one re-create a graph from above using this summary dataset <code>ds</code>?

### Answer from Comment 1
If we want to leave <code>NA</code> in the calculation of the total:
```{r}
ds<- dsN %.%
#   dplyr::filter(!is.na(attendF)) %.%   # comment out to count NA in the total
  dplyr::group_by(yearF,attendF) %.%
  dplyr::summarize(count = length( attendF)) %.%
  dplyr::mutate(total = sum(count),
              percent= count/total)
head(ds,10)
```
Missing values are used in the calculation of total responses to show the natural attrition in the study. 

```{r}
p<- ggplot( ds, aes(x=yearF, y=percent, fill=attendF))  #  keeping NA in calculations
p<- p+ geom_bar(position="stack", stat="identity")
p<- p+ scale_fill_manual(values = attcol8,
                         name="Response category" )
p<- p+ scale_y_continuous("Prevalence: proportion of total",
                          limits=c(0, 1),
                          breaks=c(.1,.2,.3,.4,.5,.6,.7,.8,.9,1))
p<- p+ scale_x_discrete("Waves of measurement",
                        limits=as.character(c(2000:2005)))
p<- p+ labs(title=paste0("In the past year, how often have you attended a worship service?"))
p

```
However, assumming that attrition is not significantly associated with the outcome measure, it would be of interest to see how relative prevalence of response endorsement changes over time, or perhaps stays in an equilibrium. For this we need to remove missing values from the calculation of the total of responses:  
```{r}
ds<- dsN %.%
  dplyr::filter(!is.na(attendF)) %.%   # comment out to count NA in the total
  dplyr::group_by(yearF,attendF) %.%
  dplyr::summarize(count = length( attendF)) %.%
  dplyr::mutate(total = sum(count),
              percent= count/total)
head(ds,10)
```

Graph reflects this accordingly:
```{r}
p<- ggplot( ds, aes(x=yearF, y=percent, fill=attendF))  #  keeping NA in calculations
p<- p+ geom_bar(position="stack", stat="identity")
p<- p+ scale_fill_manual(values = attcol8,
                         name="Response category" )
p<- p+ scale_y_continuous("Prevalence: proportion of total",
                          limits=c(0, 1),
                          breaks=c(.1,.2,.3,.4,.5,.6,.7,.8,.9,1))
p<- p+ scale_x_discrete("Waves of measurement",
                        limits=as.character(c(2000:2005)))
p<- p+ labs(title=paste0("In the past year, how often have you attended a worship service?"))
p

```