---
title: "Introduction to LCM"
output:
  html_document:
    css: ~/GitHub/Longitudinal_Models_of_Religiosity_NLSY97/www/css/thesis.css
    fig.retina: 2
    fig_width: 8
    theme: united
    toc: yes
  md_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: no
    fig_crop: no
    fig_width: 8
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    fig_width: 6.5
mainfont: Calibri
fontsize: 14pt
---


<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
require(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.

```

```{r set_options, echo=F, message=F}
require(knitr)
# getwd()
opts_chunk$set(
  results='show', 
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
#   fig.height = 4.8, 
#   fig.width = 6.5, 
  out.width = NULL,
  fig.path = 'figure_rmd/IntroLCM/',     
  dev = "png",
  dpi = 400
)
echoChunks <- FALSE
warningChunks<- FALSE
options(width=120) #So the output is 50% wider than the default.
read_chunk("./Models/Descriptives/Descriptives.R") # the file to which knitr calls for the chunks
```

```{r DeclareGlobals, echo=echoChunks, message=FALSE, results='hide'}
```

```{r LoadPackages, echo=echoChunks, message=F}
```

```{r LoadData, echo=echoChunks, message=T}
# rm(list=ls(all=TRUE))  
# dsL<-readRDS(file.path(getwd(),"Data/Derived/dsL.rds"))
# source(file.path(getwd(),"Models/Descriptives/AesDefine.R"))

```

We can retrieve data for a single person using <code>id</code>
```{r, echo=T, message=T}
ds<- dsL %>% 
  dplyr::filter(id==7499,year %in% c(2000:2011)) %>% 
  dplyr::mutate(time=year-2000) %>%
  dplyr::select(id,year,attendF,attend, time) 
print(ds)
```


```{r, echo=F}

ds<- dsL %>% 
  dplyr::filter(id==7499,year %in% c(2000:2011)) %>% 
  dplyr::select(id,year,attend,attendF) %>% 
  dplyr::mutate(time=year-2000)

# canvas
p<-ggplot(ds, aes(x=time,y=attend))
# geoms
p<- p+ geom_point(aes(fill=attendF), size=3.5, shape=21)
p<- p+ geom_line(aes(group=id), alpha=.6, linetype=2, size=.5)
# scales
p<- p+ scale_fill_manual(values=attendCol8)
p<- p+ scale_y_continuous("Church attendance",
                     limits=c(0, 8),
                     breaks=c(1:8))
p<- p+ scale_x_continuous("Years since 2000",
                   limits=c(0,11),
                   breaks=c(0:11))
# annotations
p<- p + labs(title=paste0("In the past year, how often have you attended a worship service?"))
p<- p + guides(fill = guide_legend(reverse=TRUE, title="Response category")) 
# # themes
p<- p + ggplot2::theme_bw()
p<- p + ggplot2::theme_bw(base_size=baseSize)
p<- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p<- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p<- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p<- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
trajD <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
trajD # trajectory from Data
```

Alternatively, we can enter data manually, if we want to work with artificial trajectories
```{r, eval=FALSE}
    id=rep(1,12) 
    year=seq(2000,2011,1)
    attend= c(6,7,7,6,4,3,2,1,1,1,2,2)
    attendF=ordered(attend,
            levels= unique(dsL$attend,is.na=T),
            labels=levels(dsL$attendF))
    time= year - 2000
ds<-data.frame(id,year,attendF, attend, time)
print(ds)
```

```{r, echo=F, eval=F}
# # Alternatively, for posting code online
# ds <- read.table(header = TRUE, text ="
# obs  id year attend
# 1  47 2000      6
# 2  47 2001      7
# 3  47 2002      7
# 4  47 2003      6
# 5  47 2004      5
# 6  47 2005      4
# 7  47 2006      3
# 8  47 2007      2
# 9  47 2008      1
# 10 47 2009      1
# 11 47 2010      2
# 12 47 2011      3
# 
# ")
# ds<- ds %>% dplyr::mutate(time=year-2000)
# print(ds)
```

```{r, echo=F, eval=TRUE}
# # canvas
# p<-ggplot(ds, aes(x=time,y=attend))
# # geoms
# p<- p+ geom_point(aes(fill=attendF), size=3.5, shape=21)
# p<- p+ geom_line(aes(group=id), alpha=.6, linetype=2, size=.5)
# # scales
# p<- p+ scale_fill_manual(values=attendCol8)
# p<- p+ scale_y_continuous("Church attendance",
#                      limits=c(0, 8),
#                      breaks=c(1:8))
# p<- p+ scale_x_continuous("Years since 2000",
#                    limits=c(0,11),
#                    breaks=c(0:11))
# # annotations
# p<- p + labs(title=paste0("In the past year, how often have you attended a worship service?"))
# p<- p + guides(fill = guide_legend(reverse=TRUE, title="Response category")) 
# # # themes
# p<- p + ggplot2::theme_bw()
# p<- p + ggplot2::theme_bw(base_size=baseSize)
# p<- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
# p<- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
# p<- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
# p<- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
# trajM <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
# trajM # trajectory entered Manually
trajM<- trajD
```

The simplest possible model that can be fit to these data is the intercept-only model:  

\[{y_{ti}} = {\beta _0} + {\varepsilon _{ti}}\]  

which implies that the data can be represented with a straight, zero-slope line:
```{r}
ds$m0<- predict(lm(attend ~ 1, ds))
print(ds)
p<- trajM+ ggplot2::geom_line(aes(y=m0),color="black", size=.5)
p
```

Adding a fixed slope to the equation:  

\[{y_{ti}} = {\beta _0} + {\beta _1}Tim{e_t} + {\varepsilon _{ti}}\]

```{r c5}
m1<- predict(lm(attend ~ time, ds))
ds<- ds %>% dplyr::mutate(m1=m1)
print(ds)
p<-p+ geom_line(aes(y=m1),color="#1b9e77", size=.5)
p
```


Adding the curvarture to the prediction line can be accomplished by adding a quadratic term
```{r c6}
m2 <- predict(lm(attend ~ time + I(time^2),ds))
# quadratic<- predict(lm(attend ~ poly(time,2),ds))
ds<- ds %>% mutate(m2=m2)
print(ds)
p<-p+ geom_line(aes(y=m2),color="#d95f02", size=.5)
p
# p<-p+ geom_line(aes(y=cubic),color="green", size=.5)
```

One can also add a cubic term, if the second turning point is neccessary to describe the curvature
```{r c7}
m3 <- predict(lm(attend ~ time + I(time^2) + I(time^3),ds))
ds<- ds %>% mutate( m3=m3)
print(ds)
p<-p+ geom_line(aes(y=m3),color="#7570b3", size=.5)
p
```
