---
title: "Metrics"
author: "Andriy Koval"
date: "Tuesday, June 24, 2014"
output:
  pdf_document:
    fig_width: 8
    number_sections: yes
    toc: yes
    toc_depth: 3
  html_document:
    css: ~/GitHub/Longitudinal_Models_of_Religiosity_NLSY97/www/css/thesis.css
    fig.retina: 2
    fig_width: 8
    toc: yes
  word_document:
    fig_width: 6.5
  md_document:
    toc: yes
    toc_depth: 3
mainfont: Calibri
---


<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
require(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.

```

```{r set_options, echo=F, message=F}
require(knitr)
# getwd()
opts_chunk$set(
  results='show', 
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
#   fig.height = 4.8, 
#   fig.width = 6.5, 
  out.width = NULL,
  fig.path = 'figure_rmd/Attendance/',     
  dev = "png",
  dpi = 400
)
echoChunks <- FALSE
warningChunks<- FALSE
options(width=120) #So the output is 50% wider than the default.
read_chunk("./Models/Descriptives/Descriptives.R") # the file to which knitr calls for the chunks
```

```{r DeclareGlobals, echo=echoChunks, message=FALSE}
# Defined
attcol8 # color palette for attendance
baseSize # base text size
colNA # color for the NA value in graphs

```

```{r LoadPackages, echo=echoChunks, message=F}
```

```{r LoadData, echo=echoChunks, message=T}
# select only respondence in the cross-sectional sample

```

# Mapping Church Attendance in Time

## Cross-Sectional View  

The focal variable of interest is **attend**, the item measuring church attendance for the year that preceded the interview date. The questionnaire recorded the responses on the ordinal scale.  

```{r attend_2000, echo=echoChunks, message=T, fig.height=3.5,fig.cap="Figure caption test"}
ds<- dplyr::filter(dsL,year==2000,sample==1)
p<-ggplot(ds, aes(x=attendF, fill=attendF))
p <- p + geom_bar(color="black")
p <- p + scale_fill_manual(values=attcol8, na.value=colNA, guide=FALSE)
p <- p + coord_flip()
# p <- p + guides(fill = guide_legend(reverse=TRUE, title="Attendence")) #http://www.cookbook-r.com/Graphs/Legends_(ggplot2)/#reversing-the-order-of-items-in-the-legend
p <- p + labs(title="How often did you attend a place of worship in the last year (2000)?", x="Church attendance", y="Count")
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p

```

Creating frequency distributions for each of the measurement wave we have:    

```{r attend_2000_2011, echo=echoChunks,warning=F, message=F, fig.height=3.5}
ds<- dsL %.%
  dplyr::filter(sample==1) %.%   # comment out to count NA in the total
  dplyr::group_by(yearF,attendF) %.%
  dplyr::summarize(count = length( attendF)) %.%
  dplyr::mutate(total = sum(count),
              percent= count/total)
# head(ds,10)
p<- ggplot( ds, aes(x=yearF, y=percent, fill=attendF))  #  keeping NA in calculations
# geoms
p<- p+ geom_bar(position="stack", stat="identity")
# scales
p<- p+ scale_fill_manual(values = attcol8)
p<- p+ scale_y_continuous(limits=c(0, 1),
                          breaks=c(.1,.2,.3,.4,.5,.6,.7,.8,.9,1))
p<- p+ scale_x_discrete(limits=as.character(c(2000:2011)))
# annotations
p<- p + guides(fill = guide_legend(reverse=TRUE, title="Response category")) 
p<-  p+ labs(title = "In the past year, how often have you attended a worship service?",
                         x="Wave of measurement",
                         y="Proportion of total")
# themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p
```

Here, missing values are used in the calculation of total responses to show the natural attrition in the study. Assuming that attrition is not significantly associated with  the outcome measure, we can remove missing values from the calculation of the total and look at prevalence of response endorsements over time.     

```{r attend_2000_2011_na,echo=F, message=F, warning=F, fig.height=3.5}
ds<- dsL %.%
  dplyr::filter(sample==1, !is.na(attendF)) %.%   # comment out to count NA in the total
  dplyr::group_by(yearF,attendF) %.%
  dplyr::summarize(count = length( attendF)) %.%
  dplyr::mutate(total = sum(count),
              percent= count/total)
# head(ds,10)
p<- ggplot( ds, aes(x=yearF, y=percent, fill=attendF))  #  keeping NA in calculations
# geoms
p<- p+ geom_bar(position="stack", stat="identity")
# scales
p<- p+ scale_fill_manual(values = attcol8)
p<- p+ scale_y_continuous(limits=c(0, 1),
                          breaks=c(.1,.2,.3,.4,.5,.6,.7,.8,.9,1))
p<- p+ scale_x_discrete(limits=as.character(c(2000:2011)))
# annotations
p<- p + guides(fill = guide_legend(reverse=TRUE, title="Response category")) 
p<-  p+ labs(title = "In the past year, how often have you attended a worship service?",
                         x="Wave of measurement",
                         y="Proportion of total")
# themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p
```
Tracing the rate of change of prevalence in a line graph, we see more clearly which  categores increase over time (e.g. "Never"), which decline (e.g. ""About once/week), and which stay relatively stable (e.g. "About twice/month").    

```{r attend_freq_lines, echo=F, message=F, warning=F, fig.height=4.2}
ds<- dsL %.%
  dplyr::filter(sample==1, year %in% c(2000:2011), !is.na(attend)) %.%   # comment out to count NA in the total
  dplyr::group_by(yearF,attendF) %.%
  dplyr::summarize(count = length( attendF)) %.%
  dplyr::mutate(total = sum(count),
              percent= count/total)
# head(ds,10)
p<- ggplot( ds, aes(x=yearF, y=percent, color=attendF))  #  keeping NA in calculations
# geoms
p<- p+ geom_line(aes(group=attendF), stat="identity", size=.7)
p<- p+ geom_point(aes(fill=attendF), shape=21, size=3, color="black",position=position_dodge(0.2) )
# scales
p<- p+ scale_color_manual(values = attcol8,na.value=colNA)
p<- p+ scale_fill_manual(values = attcol8,na.value=colNA)
p<- p+ scale_y_continuous(limits=c(0, .4),
                          breaks=c(0,.1,.2,.3,.4))
p<- p+ scale_x_discrete(limits=as.character(c(2000:2011)))
# # annotations
p<- p + guides(fill = guide_legend(reverse=FALSE, title="Response category")) 
p<- p + guides(color = FALSE) 
p<-  p+ labs(title = "In the past year, how often have you attended a worship service?",
                         x="Wave of measurement",
                         y="Proportion of total")
# # themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p
```

Inspecting the prevalence trajectories across races.   

```{r attend_freq_lines_race, echo=F, message=F, warning=F, fig.height=7, fig.width=10}
ds<- dsL %.%
  dplyr::filter(year %in% c(2000:2011), !is.na(attend), race !=3) %.%   # comment out to count NA in the total
  dplyr::group_by(sampleF,yearF,raceF,attendF) %.%
  dplyr::summarize(count = n()) %.%
  dplyr::mutate(total = sum(count),
                percent= count/total)

p<- ggplot( ds, aes(x=yearF, y=percent, color=attendF)) 
# geoms
p<- p+ geom_line(aes(group=attendF), stat="identity", size=.4)
p<- p+ geom_point(aes(fill=attendF), shape=21, size=2, color="black",linesize=1,position=position_dodge(0.2) )
p<- p+ facet_grid(sampleF ~ raceF)
# scales
p<- p+ scale_color_manual(values = attcol8,na.value=colNA)
p<- p+ scale_fill_manual(values = attcol8,na.value=colNA)
p<- p+ scale_y_continuous(limits=c(0, .5),
                          breaks=c(0,.1,.2,.3,.4,.5))
p<- p+ scale_x_discrete(limits=as.character(c(2000:2011)))
# # annotations
p<- p + guides(fill = guide_legend(reverse=FALSE, title="Response category")) 
p<- p + guides(color = FALSE) 
p<-  p+ labs(title = "In the past year, how often have you attended a worship service?",
             x="Wave of measurement",
             y="Proportion of total")
# # themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p <- p + ggplot2::theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5, size=8))
p
```

```{r attend_race_2000, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
ds<- dsL %.%
  dplyr::filter(year %in% c(2000), !is.na(attend), race !=3) %.%   # comment out to count NA in the total
  dplyr::group_by(sampleF,yearF,raceF,attendF) %.%
  dplyr::summarize(count = n()) %.%
  dplyr::mutate(total = sum(count),
                percent= count/total)
# ds
p<- ggplot( ds, aes(x=attendF, y=percent, fill=attendF)) 
# geoms
p<- p+ geom_bar(stat="identity")
p<- p+ facet_grid(sampleF ~ raceF)
p<- p+ geom_text(aes(label=round(percent,2)), position=position_dodge(.9), vjust=-.2, size=3.5)
# scales

p<- p+ scale_fill_manual(values = attcol8,na.value=colNA)
p<- p+ scale_y_continuous(limits=c(0, .3),
                          breaks=c(0,.1,.2,.3,.4,.5))
# # annotations
p<- p + guides(fill = guide_legend(reverse=FALSE, title="Response category")) 
p<- p + guides(color = FALSE) 
p<-  p+ labs(title = "In the past year(2000), how often have you attended a worship service?",
             x="Church attendance",
             y="Proportion of total")
# # themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p <- p + ggplot2::theme(axis.text.x = element_blank())
p
```


## Longitudinal View

Graphs above shows change in the cross-sectional distribution of responses over the years. Modeling the change in these response frequencies is handled well by Markov  models.  LCM, however, works with longitudinal data, modeling the trajectory of each individual and treating attendance as a continuous outcome.   

To demonstrate mapping of individual trajectories to time, let's select a  dataset that would include personal identifyer (**id**), cohort indicator (**byear**), wave of measurement (**year**) and the focal variable of interest - worship attendance (**attend**).    

```{r, echo=T, message=F, warning=warningChunks}
ds<- dsL %>%  dplyr::filter(year %in% c(2000:2011), id==47) %>%
              dplyr:: select(id, byear, year, attend, attendF)
print(ds)
```

The view above lists attendance data for subjust with id = 47. Mapping his attendance to time we have  .  

```{r attend_line_1id,echo=echoChunks, message=F, warning=warningChunks, fig.height=3.3 }
ds<- dsL %>% dplyr::filter(id==47,year %in% c(2000:2011)) %>% dplyr::select(idF,year,attend) %>%
  mutate(time=year-2000)

p<-ggplot(ds, aes(x=time,y=attend, color=idF))
p<-p+ geom_line(size=.4)
p<-p+ geom_point(size=2)
p<-p+ scale_y_continuous("Church attendance",
                     limits=c(0, 8),
                     breaks=c(1:8))
p<-p+ scale_x_continuous("Years since 2000",
                   limits=c(0,11),
                   breaks=c(0:11))
p <- p + scale_color_manual(values=c("purple"))
# annotations
p<-p+ labs(title=paste0("In the past year, how often have you attended a worship service?"))
p <- p + guides(color = guide_legend(title="Respondents"))
# themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p
```  

where vertical dimension maps the outcome value and the horizontal maps the time. There will be a trajecory for each of the   
```{r, echo=T, message=F, warning=warningChunks}
length(unique(dsL$id))
```
subjects in total. Unless specified otherwise, only individuals from the cross-sample will be used in the model to increase external validity.   
```{r}
ds<- dsL %>% dplyr::filter(sample==1)
```


Each of such trajectories imply a story, a life scenario. Why one person grows in his religious involvement, while other declines, or never develops an interest in the first place? To demostrate how interpretations of trajectories can vary among individuals consider the following scenario.   

Attendance trajectories of subjects with **id**s  4, 25, 35, and 47 are plotted in the next graph   

```{r attend_line_4id_years,echo=echoChunks, message=F, warning=F, fig.height=3.3}
ds<- dsL %>% dplyr::filter(id %in% c(4,25,35,47),year %in% c(2000:2011)) %>% dplyr::select(idF,year,attend) %>%
  mutate(time=year-2000)
p<-ggplot(ds, aes(x=time,y=attend, color=idF))
# geoms
p<-p+ geom_line(size=.4)
p<-p+ geom_point(size=2, position=position_dodge(0.2))
# scales
p<-p+ scale_y_continuous("Church attendance",
                     limits=c(0, 8),
                     breaks=c(1:8))
p<-p+ scale_x_continuous("Years since 2000",
                   limits=c(0,11),
                   breaks=c(0:11))
p <- p + scale_color_manual(values=c("black","red","green", "purple"))
# annotations
p<-p+ labs(title=paste0("In the past year, how often have you attended a worship service?"))
p <- p + guides(color = guide_legend(title="Respondents"))
# themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p

```

The respondent  **id**=35 reported attending no worship services in any of the years, while respodent **id**=25 seemed to frequent it, indicating weekly attendance in 8 out of the 12 years. Individual **id**=47 started as a fairly regular attendee of religious services in 2000 (5= "about twice a month"), then gradually declined his involvement to nill in 2009 and on. Respondent **id**=4, on the other hand started off with a rather passive involvement, reporting  attended church only "Once or twice"  in 2000,  maintained a low level of participation throughout the years, only to surge his attendance in 2011.  Latent curve models will describe intraindividual trajectories of change, while summarizinig the interindividual similarities and trends.    

Previous research in religiousity indicated that age might be one of the primary factors explaining interindividual differences in church attendance. To examine the role of age, we change the metric of time from waves of measurement, as in the previous graph, to biological age. There are, however, a number of ways information about age was recorded.   

```{r comparing_age, warning=F }
ds<- dsL %>% dplyr::filter(id %in% c(4,25,35,47),year %in% c(2000:2011)) %>% 
  dplyr::select(idF,attend, year, byear, ageyear, agemon, ageyear) %>% 
  mutate(time=year-2000, age=(year-byear+1), ageCurrent = agemon/12)
print(ds[ds$idF==25,])
ds
```
Note that for person **id** = 25 the age was recorded as 21 years for both 2003 and 2004. However, when you examine age in months (**agemon**) you can see this is rounding issue that disappears once a more precise scale is used.  To avoid this potentially confusing peculiarity, age in years will be either calculated as (**age** = **year** - **byear** + 1)   

Plotting age, caclulated as  <code> age = year - byear + 1 </code>   

```{r attend_line_4id_age,echo=echoChunks, message=F, fig.height= 3.3 }
p<-ggplot(ds, aes(x=age,y=attend, color=idF))
# geoms
p<-p+ geom_line(size=.4)
p<-p+ geom_point(size=2, position=position_dodge(0.2))
# scales
p<-p+ scale_y_continuous("Church attendance",
                     limits=c(0, 8),
                     breaks=c(1:8))
p<-p+ scale_x_continuous("Age in years",
                   limits=c(16,31),
                   breaks=seq(from=12, to=31,by=1))
p <- p + scale_color_manual(values=c("black","red","green", "purple"))
# annotations
p<-p+ labs(title=paste0("In the past year, how often have you attended a worship service?"))
p <- p + guides(color = guide_legend(title="Respondents"))
# themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p
```



```{r}
ds<- dsL %>% dplyr::filter(id %in% c(4,25,35,47),year %in% c(2000:2011)) %>% 
  dplyr::select(idF,year,attend,agemon,ageyear) %>% 
  mutate(time=year-2000, age=ageyear)
head(ds,12)
```

Plotting age, caclulated as  <code> age = ageyear </code>  

```{r attend_line_4id_age2,echo=echoChunks, message=F, error=FALSE, warning=FALSE, fig.height=3.3  }
p<-ggplot(ds, aes(x=age,y=attend, color=idF))
# geoms
p<-p+ geom_line(size=.4)
p<-p+ geom_point(size=2, position=position_dodge(0.2))
# scales
p<-p+ scale_y_continuous("Church attendance",
                     limits=c(0, 8),
                     breaks=c(1:8))
p<-p+ scale_x_continuous("Age in years",
                   limits=c(16,31),
                   breaks=seq(from=12, to=31,by=1))
p <- p + scale_color_manual(values=c("black","red","green", "purple"))
# annotations
p<-p+ labs(title=paste0("In the past year, how often have you attended a worship service?"))
p <- p + guides(color = guide_legend(title="Respondents"))
# themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p
```

Plotting age, caclulated as  <code> age = agemon/12 </code>  

```{r}
ds<- dsL %>% dplyr::filter(id %in% c(4,25,35,47),year %in% c(2000:2011)) %>% 
  dplyr::select(idF,year,attend,agemon,ageyear,byear) %>% 
  mutate(time=year-2000, age=agemon/12)
head(ds,12)
```


```{r attend_line_4id_age3,echo=echoChunks, message=F, error=FALSE, warning=FALSE, fig.height = 3.3 }
p<-ggplot(ds, aes(x=age,y=attend, color=idF))
# geoms
p<- p + geom_line(size=.4)
p<- p + geom_point(size=2, position=position_dodge(0.2))
# scales
p<- p + scale_y_continuous("Church attendance",
                     limits=c(0, 8),
                     breaks=c(1:8))
p<- p + scale_x_continuous("Age in years",
                   limits=c(16,31),
                   breaks=seq(from=12, to=31,by=1))
p <- p + scale_color_manual(values=c("black","red","green", "purple"))
# annotations
p<-  p + labs(title=paste0("In the past year, how often have you attended a worship service?"))
p<-  p + guides(color = guide_legend(title="Respondents"))
p<-  p + geom_point(position=position_dodge(0.2) )
# themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p
```




[1]:http://statcanvas.net/thesis/databox/
[2]:http://statcanvas.net/thesis/III_methods/03_Methods.htm
[3]:https://github.com/andkov/Longitudinal_Models_of_Religiosity_NLSY97/blob/master/Data/Derive_dsL_from_Extract.md
[4]:https://github.com/andkov/Longitudinal_Models_of_Religiosity_NLSY97/blob/master/Scripts/Data/LabelingFactorLevels.R
[5]:https://github.com/andkov/Longitudinal_Models_of_Religiosity_NLSY97/blob/master/Vignettes/dplyr/Data_Manipulation_Guide.md
[fig 3.2]: ./figure_rmd/3_Methods_Figure_3_2.png 