---
title: "Descriptives"
output:
  html_document:
    css: ~/GitHub/Longitudinal_Models_of_Religiosity_NLSY97/www/css/thesis.css
    fig.retina: 2
    fig_width: 8
    theme: united
    toc: yes
  md_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: no
    fig_crop: no
    fig_width: 8
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    fig_width: 6.5
mainfont: Calibri
---



<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
require(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.

```

```{r set_options, echo=F, message=F}
require(knitr)
# getwd()
opts_chunk$set(
  results='show', 
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
#   fig.height = 4.8, 
#   fig.width = 6.5, 
  out.width = NULL,
  fig.path = 'figure_rmd/Descriptives/',     
  dev = "png",
  dpi = 400
)
echoChunks <- FALSE
warningChunks<- FALSE
options(width=120) #So the output is 50% wider than the default.
read_chunk("./Models/Descriptives/Descriptives.R") # the file to which knitr calls for the chunks
```

```{r DeclareGlobals, echo=echoChunks, message=FALSE, results='hide'}
# aesDefs
```

```{r LoadPackages, echo=echoChunks, message=F}
```

```{r LoadData, echo=echoChunks, message=T}
# vignette("introduction",package="dplyr") # <- for more info on dplyr
dsL<-readRDS("./Data/Derived/dsL.rds")
source("./Models/LCM/LCModels.R")
```


Basic descriptives reports on the selected NLSY97 items   
[back to Appendix][appendix]

## Basic demographics 
A clean dataset **dsL** contains data on

```{r}
dplyr::summarize(dsL,count=n_distinct(id))
```
respondents. Of them one (id = 467) was removed from the dataset due to abberant age score that seemed as a coding mistake. NLSY97 contains representative household sample and the oversample of racial minorities. 

```{r, warning=F}
ds<- dsL %>% 
  dplyr::group_by(sampleF) %>% 
  dplyr::summarize (count=n_distinct(id))
ds
```



```{r basic_demo, echo=F, message=F, warning=F, fig.height=4.2}
ds<- dsL %>% dplyr::filter(id %in% c(1:9000), year==1997) %>% dplyr::select(id,sampleF,sexF, raceF) %>%
  dplyr::group_by(sampleF, sexF,raceF) %>% dplyr::summarize (count= n()) 

# ds<- reshape2::melt(ds,id.vars= colnames(ds))
# ds
p<- ggplot(ds, aes(x=raceF, y=count, fill=sexF))
#geoms
p<- p+ geom_bar(stat="identity", position="dodge")
p<- p+ facet_grid(sampleF~.)
p<- p+ geom_text(aes(label=count), position=position_dodge(.9), vjust=-.2, size=3.5)
# scales
p<- p + scale_y_continuous( limits= c(0,2700))
# annotations
p<- p + guides(fill = guide_legend( title="Sex")) 
p<- p+ labs(title = "Basic demographic constitution of NLSY97",
                         x="Race",
                         y="Sampling frame")
# themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p

```

## Distribution of age variables
The age of respondents was of particular interest and was entered as a predictor of the model outcome. NSLY97 contains static and dynamic indicators of age age. Variables  byear and bmonth were recorded once in 1997 (static) and contain respondents’ birth year and birth month respectively. Two age variables were recorded continuously at each interview (dynamic): age at the time of the interview in months (**agemon**) and in years (**ageyear**).   Next graph shows how births in the NLSY97 sample  was distributed over calendric months from 1980 to 1984:

### Months of births
```{r bmonth_dist, echo=F, warning=F, fig.height=3.3}
ds<- dsL %>% dplyr::filter(year %in% c(1997)) %>% 
  dplyr::select(id,year,byear,bmonth,bmonthF,agemon,ageyear) %>%
  dplyr::group_by(bmonth, bmonthF,byear) %>%
  dplyr::summarize(count=n()) %>% 
  dplyr::mutate (born= (byear*12+bmonth)/12) %>%
  dplyr::arrange(byear, bmonth)
head(ds, 10)
p<-ggplot(ds, aes(x=factor(born), y=count, fill=factor(byear)))
#geoms 
p<-p+geom_bar(stat="identity")
p<- p + scale_fill_manual(values=byearCol5)
p<- p + scale_x_discrete(labels=ds$bmonthF)
# annotations
p<- p + guides(fill = guide_legend(reverse=TRUE, title="Birth year")) 
p<-  p+ labs(title = "Distribution of births among NLSY97 repsondents",
                         x="The month of birth",
                         y="Count")
# themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40", size=8))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p<- p + theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5))
p
```


### Age and cohort structure 
```{r agemon_dist, echo=F, warning=F, fig.height=3.3}
ds<- dsL %>% dplyr::filter(id %in% c(1:19000), year %in% c(2000)) %>% dplyr::select(id,year,byearF,attend,agemon,ageyear) %>%
  mutate(time=year-2000, age=agemon/12)
# canvas
p<- ggplot(ds, aes(x=age, fill=byearF))
p<- p + geom_bar(bin=1/12)
p<- p + scale_fill_manual(values=byearCol5)
# scales
p<- p + scale_x_continuous(breaks=seq(16,32,1))
# annotations
p<- p + guides(fill = guide_legend(reverse=TRUE, title="Birth year")) 
p<- p + labs(title = "Age and cohort structure (2000)",
                         x="Age in years. Bin = 1 month",
                         y="Count")
# themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p
```


###  Count of respondents in age bins
The amounts of respondents in in the age bin =16 is not big, so results should be interpreted with caution.
```{r attend_race_age_binsize, echo=F, message=F, warning=F, fig.height=4, fig.width=10}
ds<- dsL %.%
  dplyr::mutate(ageF = as.factor(round(agemon/12 ))) %>%
  dplyr::filter(year %in% c(2000:2011), !is.na(attend), race !=3) %.%   
  dplyr::group_by(sampleF,raceF,ageF) %.%
  dplyr::summarize(count = n()) 
# canvas
p<- ggplot( ds, aes(x=ageF, y=count)) 
# geoms
p<- p + geom_bar(stat="identity")
p<- p+ facet_grid(sampleF~raceF)
# scales
# p<- p+ scale_fill_manual(values = c(basicDark,"white")) 
p<- p+ scale_y_continuous(limits=c(0, 4500),
                          breaks=c(1000,2000,3000,4000))
p<- p+ scale_x_discrete(limits=as.character(c(16:31)))
# # annotations
p<- p + guides(fill = guide_legend(reverse=FALSE, title="Response category")) 
p<- p + guides(color = FALSE) 
p<-  p+ labs(title = "In the past year, how often have you attended a worship service?",
             x="Wave of measurement",
             y="Proportion of total")
# # themes
p <- p + ggplot2::theme_bw()
p <- p + ggplot2::theme_bw(base_size=baseSize)
p <- p + ggplot2::theme(title=ggplot2::element_text(colour="gray20",size = 12)) 
p <- p + ggplot2::theme(axis.text=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(axis.title=ggplot2::element_text(colour="gray40"))
p <- p + ggplot2::theme(panel.border = ggplot2::element_rect(colour="gray80"))
p <- p + ggplot2::theme(axis.ticks.length = grid::unit(0, "cm"))
p <- p + ggplot2::theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5, size=8))
p
```

```{r}
ds<- dsL %.%
  dplyr::mutate(ageF = as.factor(round(agemon/12 ))) %>%
  dplyr::filter(year %in% c(2000:2011), !is.na(attend),ageF %in% c(16,32),race !=3) %.%   
  dplyr::group_by(sampleF,raceF) %.%
  dplyr::summarize(count = n()) 
head(ds, 10)

```


## Counts for data used in models
The following tables inspect the counts of valid responses (church attendance) across two time metircs (age and NLSY97 round) and cohort membership. 
```{r}
ds<- dsL %>%  # chose conditions to apply in creating dataset for modeling
  dplyr::filter(id %in% c(1:9022)) %.% # 1:9022
  dplyr::filter(year %in% c(2000:2011)) %.% # 1997:2011
  dplyr::filter(sample %in% c(0,1)) %.% # 0-Oversample; 1-Cross-Sectional
  dplyr::filter(race %in% c(1:4)) %.% # 1-Black; 2-Hispanis; 3-Mixed; 4-White
  dplyr::filter(byear %in% c(1980:1984)) %>% # birth year 1980:1984
  dplyr::mutate(age=round(agemon/12,0)) # define bin for age
length(unique(ds$id)) # total No. of respondents in dataset
``` 

```{r}
d<- ds %>% 
  filter(!is.na(attend)) # only nonmissing datapoints
table(d$byear,d$year) # Number of respondent in each NLSY97 round 
table(d$byear, d$age) # Number of respondent in each age bin 
table(d$age,d$year) # cross-section of age and round of NLSY97
```

When subsetting data to run throught the sequence of models, it would of use to see such counts. Insufficient number of observation per cell might be behind poor model estimation. For example, male blacks from the oversample may not be a good subset to fit a complex model. 
```{r}
ds<- dsL %>%  # chose conditions to apply in creating dataset for modeling
  dplyr::filter(id %in% c(1:9022)) %.% # 1:9022
  dplyr::filter(year %in% c(2000:2011)) %.% # 1997:2011
  dplyr::filter(sample %in% c(0)) %.% # 0-Oversample; 1-Cross-Sectional
  dplyr::filter(race %in% c(1)) %.% # 1-Black; 2-Hispanis; 3-Mixed; 4-White
  dplyr::filter(byear %in% c(1980:1984)) %>% # birth year 1980:1984
  dplyr::filter(sex==1) %>% # 1-Male; 2-Female
  dplyr::mutate(age=round(agemon/12,0)) # define bin for age

length(unique(ds$id)) # total No. of respondents in dataset
``` 

```{r}
d<- ds %>% 
  filter(!is.na(attend)) # only nonmissing datapoints
table(d$byear,d$year) # Number of respondent in each NLSY97 round 
table(d$byear, d$age) # Number of respondent in each age bin 
table(d$age,d$year) # cross-section of age and round of NLSY97
```


## Read more 

in <code>./Models/Descriptives</code>:   

+ [Metrics][metrics] - how values of items are labeled       
+ [Descriptives][descriptives] - basic stats of various items       
+ [Attendance][attend] - focus on church attendence over time (__Continue__)      
+ [Databox][databox]   

See also   

+ [Deriving Data from NLYS97 extract][derive]
+ [Data Manipulation Guide][manipulate]


```{r child, child = 'Citations.Rmd'}

```



