---
title: "Attend over Ages"
output:
  html_document:
    css: ~/GitHub/Longitudinal_Models_of_Religiosity_NLSY97/www/css/thesis.css
    fig.retina: 2
    fig_width: 8
    theme: united
    toc: yes
  md_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: no
    fig_crop: no
    fig_width: 8
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    fig_width: 6.5
mainfont: Calibri
---


<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
require(knitr)
# IIMPORTANT:  spell the path to the root directory
opts_knit$set(root.dir='../../../')  #Don't combine this call with any other chunk -especially one that uses file paths.

```

```{r set_options, echo=F, message=F}
require(knitr)

opts_chunk$set(
  results='show', 
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
#   fig.height = 4.8, 
#   fig.width = 6.5, 
  out.width = NULL,
  fig.path = 'Agemon_Ages/',     
  dev = "png",
  dpi = 100
)
echoChunks <- FALSE
warningChunks<- FALSE
options(width=120) #So the output is 50% wider than the default.
read_chunk("./Models/Descriptives/Descriptives.R") # the file to which knitr calls for the chunks
```

```{r DeclareGlobals, echo=echoChunks, message=FALSE, results='hide'}
aesDefs
```

```{r LoadPackages, echo=echoChunks, message=F}
```

```{r LoadData, echo=echoChunks, message=T}
# locate the graph that you would like to replicate across time
# commonGraphAge<-file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_counts.R")
# manually find and replace "agemon_" into the chunk/image name you want
# don't forget to change the output folder in fig.path = 'figure_rmd/', where images for this focal variable(s) will be stored
```
Among the respondents who were +/- 6 months from a given age.
For example, a 16 year-old  is defined as an individual between 15.5 and 16.5 years of age at the time of the interview.

## Attendance among races at age 16
```{r agemon_16, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 16
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 17
```{r agemon_17, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 17
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 18
```{r agemon_18, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 18
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 19
```{r agemon_19, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 19
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 20
```{r agemon_20, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 20
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 21
```{r agemon_21, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 21
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 22
```{r agemon_22, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 22
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 23
```{r agemon_23, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 23
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 24
```{r agemon_24, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 24
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 25
```{r agemon_25, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 25
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 26
```{r agemon_26, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 26
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 27
```{r agemon_27, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 27
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 28
```{r agemon_28, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 28
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 29
```{r agemon_29, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 29
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

## Attendance among races at age 30
```{r agemon_30, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 30
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```


## Attendance among races at age 31
```{r agemon_31, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 31
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```


## Attendance among races at age 32
```{r agemon_32, echo=F, message=F, warning=F, fig.height=2.4, fig.width=10}
t<- 32
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_agemon_ages.R"))
p
```

.   
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
# Read more 

in <code>./Models/Descriptives</code>:   

+ [Metrics][metrics] - how values of items are labeled       
+ [Descriptives][descriptives] - basic stats of various items       
+ [Attendance][attend] - focus on church attendence over time      
+ [Databox][databox]        

See also   

+ [Deriving Data from NLYS97 extract][derive]
+ [Data Manipulation Guide][manipulate]


```{r child, child = '../Citations.Rmd'}

```

