---
title: "Bars over Year"
output:
  html_document:
    css: ~/GitHub/Longitudinal_Models_of_Religiosity_NLSY97/www/css/thesis.css
    fig.retina: 2
    fig_width: 8
    theme: united
    toc: yes
    keep_md: false
  md_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: no
    fig_crop: no
    fig_width: 8
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    fig_width: 6.5
mainfont: Calibri
---

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
require(knitr)
# IIMPORTANT:  spell the path to the root directory
opts_knit$set(root.dir='../../../')  #Don't combine this call with any other chunk -especially one that uses file paths.

```

```{r set_options, echo=F, message=F}
opts_chunk$set(
  results='show', 
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
#   fig.height = 4.8, 
#   fig.width = 6.5, 
  out.width = NULL,
  fig.path = 'Attend_Race_Years/',     
  dev = "png",
  dpi = 100
)
echoChunks <- FALSE
warningChunks<- FALSE
options(width=120) #So the output is 50% wider than the default.
read_chunk("./Models/Descriptives/Descriptives.R") # the file to which knitr calls for the chunks
```

```{r DeclareGlobals, echo=echoChunks, message=FALSE, results='hide'}
aesDefs
```

```{r LoadPackages, echo=echoChunks, message=F}

```

```{r LoadData, echo=echoChunks, message=T}
# locate the graph that you would like to replicate across time  # to do: fix likes in sources
# commonGraphYears<-file.path(getwd(),
#                             "Models/Descriptives/temporal_animations/commonGraph_RaceYears.R")
# manually find and replace "attend_race_" into the chunk/image name you want
# don't forget to change the output folder in fig.path = 'figure_rmd/', where images for this focal variable(s) will be stored
```


## Attendance among races in 2000 
```{r attend_race_2000, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
t<- 2000
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_years.R"))
p
```

## Attendance among races in 2001
```{r attend_race_2001, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
t<- 2001
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_years.R"))
p
```

## Attendance among races in 2002
```{r attend_race_2002, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
t<- 2002
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_years.R"))
p
```

## Attendance among races in 2003
```{r attend_race_2003, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
t<- 2003
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_years.R"))
p
```

## Attendance among races in 2004
```{r attend_race_2004, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
t<- 2004
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_years.R"))
p
```

## Attendance among races in 2005
```{r attend_race_2005, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
t<- 2005
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_years.R"))
p
```

## Attendance among races in 2006 
```{r attend_race_2006, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
t<- 2006
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_years.R"))
p
```

## Attendance among races in 2007
```{r attend_race_2007, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
t<- 2007
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_years.R"))
p
```

## Attendance among races in 2008
```{r attend_race_2008, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
t<- 2008
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_years.R"))
p
```

## Attendance among races in 2009
```{r attend_race_2009, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
t<- 2009
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_years.R"))
p
```

## Attendance among races in 2010
```{r attend_race_2010, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
t<- 2010
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_years.R"))
p
```

## Attendance among races in 2011
```{r attend_race_2011, echo=F, message=F, warning=F, fig.height=5, fig.width=10}
t<- 2011
source(file.path(getwd(),"Models/Descriptives/temporal_animations/commonGraph_race_years.R"))
p
```


.   
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
.    
# Read more 

in <code>./Models/Descriptives</code>:   

+ [Metrics][metrics] - how values of items are labeled       
+ [Descriptives][descriptives] - basic stats of various items       
+ [Attendance][attend] - focus on church attendence over time      
+ [Databox][databox]        

See also   

+ [Deriving Data from NLYS97 extract][derive]
+ [Data Manipulation Guide][manipulate]


```{r child, child = '../Citations.Rmd'}

```

